---
title: "Analysis of adaptation experiment"
author: "Sebastian Schuster"
date: "11/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
library(gridExtra)
library(DescTools)
library(splines)
source("helpers.R")

theme_set(theme_bw())
colscale = scale_color_manual(values=c("#7CB637","#4381C1", "#666666")) 

```

```{r data, echo=FALSE}
# Load data

# Might-biased condition
trials = read.csv("../data/ling-145-data-adaptation-anon.csv")
exp_trials = read.csv("../data/ling-145-data-adaptation-exp-anon.csv")

get_correct_catch_trial_counts = function (data) {
  ret = data %>% 
  dplyr::filter(., catch_trial == 1) %>%
  group_by(subjectid) %>%
  summarise(catch_perf = sum(catch_trial_answer_correct))
  
  return(ret)
}

EXCLUDE_BELOW = 12

catch_trial_perf.trials = get_correct_catch_trial_counts(trials)
catch_trial_perf.exp_trials = get_correct_catch_trial_counts(exp_trials)
catch_trial_perf.all = rbind(catch_trial_perf.trials, catch_trial_perf.exp_trials) %>%
  group_by(subjectid) %>%
  summarise(catch_perf = sum(catch_perf))

trials = prepare_data(trials)
confident_m_speakers = trials[trials$condition == "confident_m",]$speaker
confident_m_speakers = ((as.integer(confident_m_speakers)-1)*-1) + 2
confident_m_speakers = factor(confident_m_speakers, levels=c(1, 2), labels=c("f", "m"))
trials[trials$condition == "confident_m",]$speaker = confident_m_speakers

trials$speaker = factor(trials$speaker, labels=c("confident", "cautious"))

trials = trials %>% dplyr::filter(subjectid != "12284d6642bcee215090ab80c61119fe584abc5178ed")

```

Let's first plot the mean utterance ratings before and after the exposure phase:

```{r summary_plots, fig.width=10, echo=FALSE}
trials %>% 
  mutate(phase = factor(phase, levels=c("pre_exposure", "post_exposure"), labels=c("pre-exposure", "post-exposure"), ordered = TRUE)) %>%
  group_by(phase,percentage_blue,modal,speaker) %>%
  summarize(mu = mean(rating), ci_low = ci.low(rating), ci_high = ci.high(rating)) %>%
  ggplot( aes(x=percentage_blue, y=mu, col=modal, lty=speaker)) + geom_line() +
  facet_wrap(~phase) + 
  geom_errorbar(aes(ymin = mu - ci_low, ymax = mu + ci_high)) + 
  colscale +
  xlab("event probability") +
  ylab("mean utterance rating") +
  guides(col=guide_legend(title="Expression")) +
  guides(lty=guide_legend(title="Speaker type"))  +
  theme(legend.position = "bottom")
```

### Statistical tests:

There seems to be a small numerical difference between the ratings of the _cautious_ and the _confident_ speaker. We can perform a t-test following the analysis of Yildirim et al. (2016), to determine whether this difference is significant.

```{r test, echo=FALSE}

auc_method1 = function(rating_m, percentage_blue) {
      #estimate linear model
    d= data.frame(rating_m = rating_m, percentage_blue=percentage_blue)
    model = lm(formula = rating_m ~ ns(percentage_blue, df = 4), data = d)
  
    pred = data.frame(0:1000)
    pred$percentage_blue = (0:1000)/10
    pred$rating = predict(model, pred)

    auc = AUC(x = pred$percentage_blue, y=pred$rating)
    return(auc)
}

auc_for_participants = function(d, method) {
  
  aucs = data.frame()

  #i = 1
  
  for (wid in unique(d$subjectid)) {
    d.might_ratings = d %>% 
      filter (., subjectid == wid) %>%
      filter (modal == "might") %>%
      group_by(subjectid, percentage_blue, speaker) %>%
      summarise(rating_m = mean(rating)) %>%
      group_by(subjectid, speaker) %>%
      summarise(auc_might = method(rating_m, percentage_blue))
    

    d.probably_ratings = d %>% 
      filter (., subjectid == wid) %>%
      filter (modal == "probably") %>%
      group_by(subjectid, percentage_blue, speaker) %>%
      summarise(rating_m = mean(rating)) %>%
      group_by(subjectid, speaker) %>%
      summarise(auc_probably = method(rating_m, percentage_blue))
    
   
    aucs = rbind(aucs,  merge(d.might_ratings, d.probably_ratings))
    #aucs$auc_probably[i] = method(d.probably_ratings)

    #i = i + 1
  }
  
  aucs$auc_diff = aucs$auc_might - aucs$auc_probably
  
  return(aucs)
}

aucs = auc_for_participants(trials, method=auc_method1)

aucs.all = aucs %>% 
  group_by(., speaker) %>% 
  summarise(., auc_diff_m = mean(auc_diff), 
               ci_high = ci.high(auc_diff), 
               ci_low = ci.low(auc_diff))




ggplot(aucs.all, aes(x=0, y=auc_diff_m, color=speaker)) +
    geom_errorbar(aes(ymin=auc_diff_m-ci_low, ymax=auc_diff_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("AUC difference (might ratings - probably ratings)") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) 

```

```{r}
print(summary(lm(auc_diff ~ speaker, data = aucs)))
```

Both the plot and the t-test suggest, that the manipulation did not have an effect and that there is no significant difference between the ratings for the _cautious_ and the _confident_ speaker.




