---
title: "Theta posteriors"
author: "Sebastian Schuster"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rwebppl)
library(ggplot2)
library(tidyr)
library(dplyr)
library(data.table)
library(gridExtra)


```


```{r load_data} 

load_data_for_plotting = function(fname) {


d = read.csv(fname)

d$modal1 = gsub('"', '', d$modal1)
d$modal2 = gsub('"', '', d$modal2)
d$pair = gsub('"', '', d$pair)
d$color = gsub('"', '', d$color)


drops <- c("modal1","rating1")
d2 = d[ , !(names(d) %in% drops)]
setnames(d2, old=c("rating2","modal2"), new=c("rating", "modal"))

drops <- c("modal2","rating2")
d3 = d[ , !(names(d) %in% drops)]
setnames(d3, old=c("rating1","modal1"), new=c("rating", "modal"))

drops <- c("modal2", "rating2", "modal1", "rating1")
d4 = d[ , !(names(d) %in% drops)]
d4$rating = d4$rating_other
d4$modal = "other"

d = rbind(d2, d3, d4)

d$modal = factor(d$modal)

d$percentage_blue_f = factor(d$percentage_blue)


d_blue = d %>% filter(., grepl("blue", sentence2))
d_orange = d %>% filter(., grepl("orange", sentence2))

#ggplot(d_orange, aes(x=percentage_blue, y=rating)) + geom_point(aes(col=modal)) +
# geom_smooth(aes(col=modal))

#ggplot(d_blue, aes(x=percentage_blue, y=rating)) + geom_point(aes(col=modal)) +  geom_smooth(aes(col=modal))

d_orange_reverse = d_orange
d_orange_reverse$percentage_blue = 100-d_orange$percentage_blue

d_comparison = rbind(d_blue, d_orange_reverse)
d_comparison$blue= grepl("blue", d_comparison$sentence2)
d_comparison$percentage_blue_f = factor(d_comparison$percentage_blue)

return(d_comparison)

}
  
load_data = function(fname) {
    d = read.csv(fname)
  
  d$modal1 = gsub('"', '', d$modal1)
  d$modal2 = gsub('"', '', d$modal2)
  d$pair = gsub('"', '', d$pair)
  d$color = gsub('"', '', d$color)

  
  d_blue = d %>% filter(., grepl("blue", sentence2))
  d_orange = d %>% filter(., grepl("orange", sentence2))
  
  #ggplot(d_orange, aes(x=percentage_blue, y=rating)) + geom_point(aes(col=modal)) +
  # geom_smooth(aes(col=modal))
  
  #ggplot(d_blue, aes(x=percentage_blue, y=rating)) + geom_point(aes(col=modal)) +    geom_smooth(aes(col=modal))
  
  d_orange_reverse = d_orange
  d_orange_reverse$percentage_blue = 100-d_orange$percentage_blue
  
  d_comparison = rbind(d_blue, d_orange_reverse)
  d_comparison$blue= grepl("blue", d_comparison$sentence2)
  
  return(d_comparison)
}

make_plots = function(d) {
  p1 = ggplot(d, aes(x=percentage_blue, y=rating)) + geom_point(aes(col=modal)) +  geom_smooth(aes(col=modal), method="loess") + ggtitle(d$pair[1]) + xlab("percentage")

p2 = ggplot(d, aes(x=percentage_blue_f, y=rating, fill=modal)) + 
  geom_boxplot() +
  ggtitle(d$pair[1]) + xlab("percentage")


return(list("p1" = p1, "p2" = p2))
}


```


```{r fig1, fig.width=10}

d = load_data("../../experiments/0_pre_test/data/0_pre_test-cond5_2-trials.csv")

d2 = load_data_for_plotting("../../experiments/0_pre_test/data/0_pre_test-cond5_2-trials.csv")



ps = make_plots(d2)

grid.arrange(ps$p1, ps$p2, ncol=2)


```



```{r model}

webppl_model = '

var modal1 = data.modal1[0];
var modal2 = data.modal2[0];



var limitedDiscreteUniform = cache(function(a,b) {
return Infer({method: "enumerate"}, function(){
  var bins = _.range(a, b + 1 , 5);
  //var actualBins =  _.range(a, b + 1 , 5);
  //var actualBinCounts = actualBins.length;
  //var fakeBinCounts = bins.length - actualBinCounts;
  //var remainingProb = a > 0 ? 0.02 * fakeBinCounts : 0;
  var probs = map(function(x){
  //  if (x < a) {
  //    return remainingProb / fakeBinCounts;
  //  } else {
  //    return (1-remainingProb) / actualBinCounts;
    return 1 / bins.length;
  //  }
} , bins);
  return bins[discrete(probs)];
});
});

var discreteUniform = limitedDiscreteUniform(0,100);

var discreteBeta = cache(function(params) {
  return Infer({method: "enumerate"}, function(){
    var bins = _.range(0, 1.01, .05);
    var probs = map(function(b){ Math.exp(Beta(params).score(b)) } , bins);
    var ret = bins[discrete(probs)];
    return ret;
  }
)});

var thetaParamDist = Infer({method: "enumerate"}, function(){
  var bins = _.range(0.1, 30.1 , 0.5);
  var probs = map(function(x){
      return 1 / bins.length;
    } , bins);
    var ret = bins[discrete(probs)];
  return ret;
});


var thetaParamPrior = function() {
return sample(Uniform({a:0, b:30}));  
//return sample(thetaParamDist);
}

var costPrior = function() {
  return sample(Uniform({a:0, b:5})); 
}

var ratPrior = function() {
    return sample(Uniform({a:0.1, b:3})); 
}


var thetaPrior = function() {
    return sample(discreteUniform)
}


var probPrior = function() {
    return sample(discreteUniform)
}

  var utterances = (function() {
   if (modal1 == "bare" || modal2 == "bare") {
       return ["bare not", modal1, modal2, "other"];
    } else {
     return ["bare not", modal1, modal2, "bare", "other"];
   }
})();



var utterancePrior = function() {
  return uniformDraw(utterances);
}

var literalListener = function(utterance, thetas) {
  if (flip(0.05)) {
    return discreteUniform; //to account for noise
  } else if (utterance == "bare not") {
    return limitedDiscreteUniform(0, 100-thetas["bare"]);
  } else if (utterance == "other") {
        return discreteUniform;
  } else if (utterance == modal1) {
    return limitedDiscreteUniform(thetas["modal1"], 100);
  } else if (utterance == modal2) {
    return limitedDiscreteUniform(thetas["modal2"], 100);
  } else {
    return limitedDiscreteUniform(thetas[utterance], 100);
  }
};


var thetas_tmp = {might: 50, bare: 90};
var pragmaticSpeaker =  cache(function(prob, alphas, betas, costs, rat_alpha) {
  var theta_modal1_dist = discreteBeta({a: alphas["modal1"], b: betas["modal1"]});
  var theta_modal2_dist = discreteBeta({a: alphas["modal2"], b: betas["modal2"]});
  var theta_bare_dist = modal1 != "bare" && modal2 != "bare" ? discreteBeta({a: data.alpha_bare[0], b: data.beta_bare[0]}): null;
  return Infer({model: function() {
    var utt = utterancePrior();
    var cost = costs[utt];
      var theta_modal1 = (utt == modal1 || (utt == "bare not" && modal1 == "bare")) ? Math.round(sample(theta_modal1_dist) * 100) : 0;
      var theta_modal2 = (utt == modal2 || (utt == "bare not" && modal2 == "bare")) ? Math.round(sample(theta_modal2_dist) * 100) : 0;
      var theta_bare =  modal1 == "bare" ? theta_modal1 : (modal2 == "bare" ? theta_modal2 : ((utt == "bare" || utt == "bare not") ? Math.round(sample(theta_bare_dist) * 100) : 0 )); 
      var thetas = {
        modal1: theta_modal1,
        modal2: theta_modal2,
        bare: theta_bare
      }
    factor(rat_alpha * (literalListener(utt, thetas).score(prob) - cost) ); 
    return (utt == modal1 || utt == modal2) ? utt : "other";
  }, method:"enumerate"});
});

var model = function() {
return thetaParamPosterior();
}

var thetaParamPosterior = function() {
  var alphas = {
    modal1: thetaParamPrior(),
    modal2: thetaParamPrior()
  };

  
  var betas = {
    modal1: thetaParamPrior(),
    modal2: thetaParamPrior()
  };

  var costs = {
    "bare": modal1 != "bare" && modal2 != "bare" ? costPrior() : 0,
    "might": modal1 != "might" && modal2 != "might" ? costPrior() : 0,
    "probably": modal1 != "probably" && modal2 != "probably" ? costPrior() : 0,
    "bare not": costPrior(),
    "other": costPrior(),
  };

  var rat_alpha = ratPrior();

  map(function(d) {
    var p = d.percentage_blue;
    var utt = d.modal;
    observe( pragmaticSpeaker(p, alphas, betas, costs, rat_alpha), 
      utt);  
  }, data.obs);
  return {alpha_modal1: alphas["modal1"], alpha_modal2: alphas["modal2"], 
          beta_modal1: betas["modal1"], beta_modal2: betas["modal2"],
          cost_bare: costs["bare"], cost_might: costs["might"],
          cost_probably: costs["probably"], cost_bare_not: costs["bare not"],
          cost_other: costs["other"], rat_alpha: rat_alpha};
}

'





```


```{r call}

  modals = c("bare", "might", "probably", "could", "looks_like", "think", "probably not", "bare not")



estimate_model = function(d) {
  
  drops <- c("sentence1", "sentence2")

  d_obs = d[ , !(names(d) %in% drops)]

  d_obs =  do.call("rbind", replicate(10, d_obs, simplify = FALSE))

  print(nrow(d_obs))
  
  d_obs = d_obs %>%
    rowwise() %>%
    mutate(rating1 = max(0, rating1), rating2 = max(0, rating2), rating_other = max(0, rating_other))
  d_obs = d_obs %>%
    rowwise() %>%
    mutate(modal = sample(c(modal1, modal2, "other" ), prob = c(rating1, rating2, rating_other), size=1) )  

 # d_obs = d_obs %>% filter(., percentage_blue > 0)
  
  print(head(d_obs))
  
  data_from_R = list(obs = d_obs, modal1=d_obs$modal1[1], modal2=d_obs$modal2[1], alpha_bare=7.58, beta_bare=0.0329)
    maxIter = 10000

  post <- webppl(program_code = webppl_model, model_var = "model",
    inference_opts = list(method = "MCMC", burn=500, samples=maxIter, verbose=TRUE),
    data_var = "data", data = data_from_R)
  
  str(post)
  
  print(post)
  print(post %>% filter(., Iteration == maxIter))
  print(post %>% group_by (., Parameter) %>% summarise(., val = mean(value)))

  #  print(max(post$prob))

  #  print(sum(post$prob))
  
    
 # theta_might_alpha_dist = post %>% 
 #   group_by(., alpha_modal1) %>% 
 #   summarise(., marginal = sum(prob))

# #theta_might_beta_dist = post %>% 
#   group_by(., beta_modal1) %>% 
#   summarise(., marginal = sum(prob))

# print("might alpha")
# print(theta_might_alpha_dist)
# 
# print("might beta")
# print(theta_might_beta_dist)
# 
# theta_bare_alpha_dist = post %>% 
#   group_by(., alpha_modal2) %>% 
#   summarise(., marginal = sum(prob))

# 
# theta_bare_beta_dist = post %>% 
#   group_by(., beta_modal2) %>% 
#   summarise(., marginal = sum(prob))
# 
# print("bare alpha")
# print(theta_bare_alpha_dist)

# print("bare beta")
# print(theta_bare_beta_dist)

 # colnames(df) <- seq(0,1, 0.001)
 # df$modal = modals
 # 
 # d = df %>% gather(., prob, score,-modal)
 # d[!(d$modal %in% current_modals),]$modal = "other"
 # 
 # lvls = current_modals
 # lvls[3] = "other"
 # 
 # d$modal = factor(d$modal, levels = lvls, ordered = TRUE)
 # 
 # d$prob = as.double(d$prob)
 # 
 # d = d %>% group_by(., prob, modal) %>% summarise(., score=sum(score))
 # 
 # p = ggplot(d, aes(x=prob, y=score, color=modal)) + geom_smooth(method="loess")
 # 
 # return(p)
}





```


```{r}

  estimate_model(d)



```



## Modeling assumptions

* the model:

$L(\phi | utt, \theta) \propto 1[\phi > \theta_utt]$ (with special cases for "bare not" and "other")

$S(utt | \phi) = \integral_0^1 P(\theta) exp(\alpha * (log L(\phi | utt, \theta) - C(utt))$$



 * The threshold theta for each of the modals is sampled from a Beta distribution with two parameters, which are estimated from the data
 * alpha = 2
 * possible utterances: "bare not", modal1, modal2, "bare", other
 * everything that is not modal1, modal2 is counted as "other" for inferring from the data
 * "bare not" is the inverse of "bare", i.e., [[bare not p]] = 1 if P(p) < theta_bare
 * "other" is uniformly distributed
 * "other" utterances have a cost of 1
 * The speaker marginalizes over the distributions of theta
 * (Could also be the listener?)
 * We assume uniform priors over probability of getting a blue one
 * We introduce error by having the listener sample from a uniform distribution over certainties with probability 0.05 (TODO: perhaps tweak?)
 * from each speaker "distribution"" from the data, we sample 10 points to estimate parameters
 
 * for the versions without ratings for bare utterances, we specify alpha and beta for theta_bare
 
 * Next steps:
   * verify if this actually works for all modals
   * tweak all parameters
   * model adaptation as a change in the prior of thetas
   * verify model predictions by computing R^2
   * 

 * What to do about Cleo's comment?
 
 * figure out whether gender effect is significant ==> yes, for probably-biased condition; no for might-biased condition
 
 * also run study 
 

