---
title: "Analysis of comprehension prior experiment"
author: "Sebastian Schuster"
date: "4/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(lme4)
library(lmerTest)

source("../../1_adaptation/analysis/helpers.R")

```

```{r utils, include=FALSE}
remove_quotes = function(d) {
  d$modal = gsub('"', '', d$modal)
  return(d)
}

computeMeans = function(d) {
  d.means = d %>% 
    mutate(weighted_p = percentage_blue / 100 * rating_norm, weighted_p2 =  (percentage_blue / 100)^2 * rating_norm) %>% 
    group_by(workerid, modal, cond) %>% 
    summarize(mu=sum(weighted_p)/sum(rating_norm), 
              sd=sqrt(sum(weighted_p2)/sum(rating_norm) - (sum(weighted_p)/sum(rating_norm))^2))
    
  return(d.means)
  
}


```

```{r load_data, echo=FALSE}

trials.might = remove_quotes(read.csv("../data/3_comprehension_prior-might-trials.csv"))
trials.probably = remove_quotes(read.csv("../data/3_comprehension_prior-probably-trials.csv"))


trials.might = trials.might %>% group_by(workerid) %>% mutate(rating_norm = rating / max(sum(rating),0.001), cond = "might") %>% ungroup()
trials.probably = trials.probably %>% group_by(workerid) %>% mutate(rating_norm = rating / max(sum(rating), 0.001),  cond = "probably") %>% ungroup()

max_workerid = max(trials.might$workerid) + 1

trials.probably = trials.probably %>% mutate(workerid = workerid + max_workerid)

exp_trials.might = read.csv("../data/3_comprehension_prior-might-exp_trials.csv")
exp_trials.probably = read.csv("../data/3_comprehension_prior-probably-exp_trials.csv")

exp_trials.might$percentage_blue =  as.numeric(as.character(gsub('"', '', exp_trials.might$percentage_blue)))
exp_trials.probably$percentage_blue =  as.numeric(as.character(gsub('"', '', exp_trials.probably$percentage_blue)))

exp_trials.probably = exp_trials.probably %>% mutate(workerid = workerid + max_workerid)

exp_trials_counts.might = exp_trials.might %>% group_by(workerid, percentage_blue) %>% summarise(exp_count = n()) %>% group_by(workerid) %>% mutate(exp_rel_count = exp_count / sum(exp_count))
exp_trials_counts.probably = exp_trials.probably %>% group_by(workerid, percentage_blue) %>% summarise(exp_count = n()) %>% group_by(workerid) %>% mutate(exp_rel_count = exp_count / sum(exp_count))

exp_trials_counts.might$cond = "exposure"
exp_trials_counts.probably$cond = "exposure"


```

## Catch trial performance

```{r catch_trials, echo=FALSE}

get_correct_catch_trial_counts = function (data) {
  ret = data %>% 
  filter(., catch_trial == 1) %>%
  group_by(workerid) %>%
  summarise(catch_perf = sum(catch_trial_answer_correct))
  
  return(ret)
}

EXCLUDE_BELOW = 3

catch_trial_perf.all.might = get_correct_catch_trial_counts(exp_trials.might)


catch_trial_perf.all.probably = get_correct_catch_trial_counts(exp_trials.probably)

exclude.might = catch_trial_perf.all.might %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid

exclude.probably = catch_trial_perf.all.probably %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid

print(paste("Excluded", length(exclude.might), "participants in might-biased condition."))
print(paste("Excluded", length(exclude.probably), "participants in probably-biased condition."))


#final data
d.might = trials.might %>% filter(., !(workerid %in% exclude.might))
d.probably = trials.probably %>% filter(., !(workerid %in% exclude.probably))

d.all = rbind(d.might, d.probably)

exp_trials_counts.might = exp_trials_counts.might  %>% filter(., !(workerid %in% exclude.might))
exp_trials_counts.probably = exp_trials_counts.probably  %>% filter(., !(workerid %in% exclude.probably))
exp_trials_counts.all = rbind(exp_trials_counts.might, exp_trials_counts.probably)


```

```{r perf_plot_1, fig.height=2,fig.width=10, echo=FALSE}

cutoff <- data.frame(yintercept=EXCLUDE_BELOW-0.5, cutoff=factor(EXCLUDE_BELOW-0.5))


ggplot(catch_trial_perf.all.probably, aes(x=factor(workerid), y=catch_perf)) + geom_bar(stat = "identity") + ggtitle("Catch trial performace in probably-biased cond.") + xlab ("participant") + ylab("correct responses") + geom_hline(aes(yintercept=yintercept, linetype=cutoff, color="red"), data=cutoff) + theme(legend.position="none")

```

```{r perf_plot_2, fig.height=2,fig.width=10, echo=FALSE}

ggplot(catch_trial_perf.all.might, aes(x=factor(workerid), y=catch_perf)) + geom_bar(stat = "identity") + ggtitle("Catch trial performace in might-biased cond.") + xlab ("participant") + ylab("correct responses") + geom_hline(aes(yintercept=yintercept, linetype=cutoff, color="red"), data=cutoff) + theme(legend.position="none")
```

## Results

### Comparison between conditions (raw ratings)

```{r per_modal, echo=FALSE, fig.width=10}

d.plot = d.all
d.plot$cond = factor(d.plot$cond, labels = c("might-biased", "probably-biased"), levels=c("might", "probably"))


ggplot(d.plot, aes(y=rating, x=percentage_blue, col=cond)) + 
  geom_point() + 
  geom_smooth(method="loess") + 
  xlab("percentage blue") + 
  ylab("raw rating") +
  guides(col=guide_legend(title="Condition"))

```

### Comparison between conditions (normalized ratings)

```{r per_modal_norm, echo=FALSE, fig.width=10}


ggplot(d.plot, aes(y=rating_norm, x=percentage_blue, col=cond)) + 
  geom_point() + 
  geom_smooth(method="loess") + 
  xlab("percentage blue") + 
  ylab("normalized rating") +
  guides(col=guide_legend(title="Condition"))



ggplot(d.plot, aes(y=rating_norm, x=factor(percentage_blue), col=cond)) + 
  geom_boxplot() +
  xlab("percentage blue") + 
  ylab("normalized rating") +
  guides(col=guide_legend(title="Condition"))



```

```{r fig.height=20, fig.width=10, echo=FALSE}
ggplot(d.plot, aes(y=rating_norm, x=percentage_blue, col=cond)) + 
  geom_point() + 
  geom_line() +
  geom_line(aes(x=percentage_blue, y=exp_rel_count), data = exp_trials_counts.all) + 
  facet_wrap(~workerid) +
  xlab("percentage blue") + 
  ylab("normalized rating") +
  guides(col=guide_legend(title="Condition"))
```



### Correlation between exposure to gumball machines and ratings?

```{r echo=FALSE}

d.merged = merge(d.plot, exp_trials_counts.all, by=c("workerid", "percentage_blue"))

ggplot(d.merged, aes(x=exp_rel_count, y= rating_norm)) +
  geom_point() +
  geom_smooth(method="lm")



model = lmer(formula = rating_norm ~ exp_rel_count + (1 + exp_rel_count | workerid), data = d.merged)

summary(model)

```


```{r echo=FALSE, fig.width=10}

d.means = computeMeans(d.plot)

ggplot(d.means, aes(y=mu, x=cond)) +
  geom_violin()

ggplot(d.means, aes(y=sd, x=cond)) +
  geom_violin()

model2 = lm(data=d.means, formula = mu ~ cond)

print(summary(model2))

model3 = lm(data=d.means, formula = sd ~ cond)

print(summary(model3))

```

### Correlation for non-uniform participants?

Look only at participants whose sd is below the 25th percentile or above the 75 percentile  (i.e., their distributions are either more peaked in the middle or at the endpoints)

```{r fig.width=10}

med =  quantile(na_if(d.means$sd, NaN), probs=c(.25, .75), na.rm = T)
print(d.means$sd)
non_uniform_workers = d.means %>% filter(., sd < med[1] || sd > med[2])
print(non_uniform_workers)

d.merged.non_uniform = d.merged %>% filter(workerid %in% non_uniform_workers$workerid)
#
ggplot(d.merged.non_uniform, aes(x=exp_rel_count, y= rating_norm)) +
  geom_point() +
  geom_smooth(method="lm")



model = lm(formula = rating_norm ~ exp_rel_count, data = d.merged.non_uniform)

summary(model)
```


```{r fig.height=20, fig.width=10, echo=FALSE}
ggplot(d.plot %>% filter(workerid %in% non_uniform_workers$workerid), aes(y=rating_norm, x=percentage_blue, col=cond)) + 
  geom_point() + 
  geom_line() +
  geom_line(aes(x=percentage_blue, y=exp_rel_count), data = exp_trials_counts.all %>% filter(workerid %in% non_uniform_workers$workerid)) + 
  facet_wrap(~workerid) +
  xlab("percentage blue") + 
  ylab("normalized rating") +
  guides(col=guide_legend(title="Condition"))
```