---
title: "Analysis of comprehension experiment (with balanced exposures)"
author: "Sebastian Schuster"
date: "05/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
source("../../1_adaptation/analysis/helpers.R")

theme_set(theme_bw())

colscale = scale_color_manual(values=c("#7CB637","#4381C1", "#666666")) 
colscale_fill = scale_fill_manual(values=c("#7CB637","#4381C1", "#666666"))

```

```{r utils, include=FALSE}
remove_quotes = function(d) {
  d$modal = gsub('"', '', d$modal)
  d$color = gsub('"', '', d$color)
  return(d)
}
```

```{r load_data, echo=FALSE}

trials.might = remove_quotes(read.csv("../data/12_comprehension_coins_balanced-might-trials.csv"))
trials.probably = remove_quotes(read.csv("../data/12_comprehension_coins_balanced-probably-trials.csv"))

trials.might[trials.might$color == "orange",]$percentage_blue = 100 - trials.might[trials.might$color == "orange",]$percentage_blue
trials.probably[trials.probably$color == "orange",]$percentage_blue = 100 - trials.probably[trials.probably$color == "orange",]$percentage_blue

trials.might = trials.might %>% group_by(modal, color, workerid) %>% mutate(rating_norm = rating / sum(rating), cond = "might") %>% ungroup()
trials.probably = trials.probably %>% group_by(modal, color, workerid) %>% mutate(rating_norm = rating / sum(rating),  cond = "probably") %>% ungroup()

exp_trials.might = read.csv("../data/12_comprehension_coins_balanced-might-exp_trials.csv")
exp_trials.probably = read.csv("../data/12_comprehension_coins_balanced-probably-exp_trials.csv")

```

## Catch trial performance

```{r catch_trials, echo=FALSE}

get_correct_catch_trial_counts = function (data) {
  ret = data %>% 
  filter(., catch_trial == 1) %>%
  group_by(workerid) %>%
  summarise(catch_perf = sum(catch_trial_answer_correct))
  
  return(ret)
}

EXCLUDE_BELOW = 4

catch_trial_perf.all.might = get_correct_catch_trial_counts(exp_trials.might)


catch_trial_perf.all.probably = get_correct_catch_trial_counts(exp_trials.probably)

exclude.might = catch_trial_perf.all.might %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid

exclude.probably = catch_trial_perf.all.probably %>%
  filter(catch_perf < EXCLUDE_BELOW) %>%
  .$workerid

print(paste("Excluded", length(exclude.might), "participants in might-biased condition."))
print(paste("Excluded", length(exclude.probably), "participants in probably-biased condition."))


#final data
d.might = trials.might %>% filter(., !(workerid %in% exclude.might))
d.probably = trials.probably %>% filter(., !(workerid %in% exclude.probably))

d.all = rbind(d.might, d.probably)


```

```{r perf_plot_1, fig.height=2,fig.width=10, echo=FALSE}

cutoff <- data.frame(yintercept=EXCLUDE_BELOW-0.5, cutoff=factor(EXCLUDE_BELOW-0.5))


ggplot(catch_trial_perf.all.probably, aes(x=factor(workerid), y=catch_perf)) + geom_bar(stat = "identity") + ggtitle("Catch trial performace in probably-biased cond.") + xlab ("participant") + ylab("correct responses") + geom_hline(aes(yintercept=yintercept, linetype=cutoff, color="red"), data=cutoff) + theme(legend.position="none")

```

```{r perf_plot_2, fig.height=2,fig.width=10, echo=FALSE}

ggplot(catch_trial_perf.all.might, aes(x=factor(workerid), y=catch_perf)) + geom_bar(stat = "identity") + ggtitle("Catch trial performace in might-biased cond.") + xlab ("participant") + ylab("correct responses") + geom_hline(aes(yintercept=yintercept, linetype=cutoff, color="red"), data=cutoff) + theme(legend.position="none")
```

## Results

### Comparison between conditions for each modal (raw ratings)

```{r per_modal, echo=FALSE, fig.width=10} 

d.plot = d.all
d.plot$cond = factor(d.plot$cond, labels = c("might-biased", "probably-biased"), levels=c("might", "probably"))
d.plot$modal = factor(d.plot$modal, levels= c("might", "probably", "bare"), ordered = T)

d_means = d.plot %>%
  group_by(workerid, percentage_blue, modal, cond) %>% 
  summarise(participant_mean = mean(rating), participant_mean_norm=mean(rating_norm)) %>%
  group_by(percentage_blue, modal, cond) %>%
  summarise(mu = mean(participant_mean),
         ci_high = ci.high(participant_mean), 
         ci_low = ci.low(participant_mean),
         mu_norm = mean(participant_mean_norm),
         ci_high_norm = ci.high(participant_mean_norm),
         ci_low_norm = ci.low(participant_mean_norm))

ggplot(d_means, aes(y=mu, x=percentage_blue, col=cond)) + 
  geom_point() + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Condition")) +
  colscale +
  theme(legend.position="bottom") +
  facet_wrap(~modal)

```

### Comparison between conditions for each modal (normalized ratings)

```{r per_modal_norm, echo=FALSE, fig.width=10}
ggplot(d_means, aes(y=mu_norm, x=percentage_blue, col=cond)) + 
  geom_point() + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_errorbar(aes(ymin=mu_norm-ci_low_norm, ymax=mu_norm+ci_high_norm), width=.1) +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Condition")) +
  colscale +
  theme(legend.position="bottom") +
  facet_wrap(~modal)

```

### Comparison between modals for each condition (raw ratings)

```{r per_cond, echo=FALSE, fig.width=10} 

ggplot(d_means, aes(y=mu, x=percentage_blue, col=modal)) + 
  geom_point() + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_errorbar(aes(ymin=mu-ci_low, ymax=mu+ci_high), width=.1) +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expression")) +
  colscale +
  theme(legend.position="bottom") +
  facet_wrap(~cond)

```


### Comparison between modals for each condition (normalized ratings)


```{r per_cond_norm, echo=FALSE, fig.width=10} 


ggplot(d_means, aes(y=mu_norm, x=percentage_blue, col=modal)) + 
  geom_point() + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_errorbar(aes(ymin=mu_norm-ci_low_norm, ymax=mu_norm+ci_high_norm), width=.1) +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expression")) +
  colscale +
  theme(legend.position="bottom") +
  facet_wrap(~cond)


```

```{r per_cond_by_part, echo=FALSE, fig.width=10, fig.height=25}
ggplot( d.plot %>%
  group_by(workerid, percentage_blue, modal, cond) %>% 
  summarise(participant_mean = mean(rating_norm)) , 
    aes(x=percentage_blue, y=participant_mean, col=modal)
  ) + 
  xlab("% blue gumballs") +
  ylab("mean ratings") +
  geom_line() +
  geom_point(size=1) +
  guides(col=guide_legend(title="Expr.")) +
  colscale +
  theme(legend.position="bottom") + facet_wrap(~cond+workerid, ncol = 8)
 


```

## Tests

t-test for difference in expected value for "might" distributions:


```{r test1, echo=FALSE}

# computes the mean of the distribution for each worker and modal
computeMeans = function(d) {
  d.means = d %>% 
    mutate(weighted_p = percentage_blue / 100 * rating_norm, weighted_p2 =  (percentage_blue / 100)^2 * rating_norm) %>% 
    group_by(workerid, modal) %>% 
    summarize(mu=sum(weighted_p)/sum(rating_norm), 
              sd=sqrt(sum(weighted_p2)/sum(rating_norm) - (sum(weighted_p)/sum(rating_norm))^2))
    
  return(d.means)
  
}


means.might_biased = computeMeans(d.might)
means.probably_biased = computeMeans(d.probably)

means.might_biased$cond = "might-biased"
means.probably_biased$cond = "probably-biased"

means.all = rbind(means.might_biased, means.probably_biased)
means.all = means.all %>% 
  group_by(cond, modal) %>% 
  summarise(mu_m = mean(mu), 
               ci_high = ci.high(mu), 
               ci_low = ci.low(mu))


t.test(means.might_biased %>% filter(modal == "might") %>% .$mu, means.probably_biased %>% filter(modal == "might") %>% .$mu, var.equal = TRUE)

```

t-test for difference in expected value for "probably" distributions:

```{r test2, echo=FALSE, fig.width=6}
t.test(means.might_biased %>% filter(modal == "probably") %>% .$mu, means.probably_biased %>% filter(modal == "probably") %>% .$mu, var.equal = TRUE)

ggplot(means.all %>% filter(modal != "bare"), aes(x=0, y=mu_m, color=cond)) +
    geom_errorbar(aes(ymin=mu_m-ci_low, ymax=mu_m+ci_high), width=.1) +
    geom_point() +
    xlab("") +
    ylab("expected value of distribution") +
    theme(axis.ticks=element_blank(), axis.text.x=element_blank(),
          panel.grid.minor=element_blank(), 
          plot.background=element_blank()) +
    xlim(-.2, .2) +
    facet_wrap(~modal) +
    guides(col=guide_legend(title="Condition"))


```

  
